### 2012-09-16 ProjectProposal/Project.tex
Proposal

\documentclass[a4paper,10pt]{article}
%\documentclass[a4paper,10pt]{scrartcl}

\usepackage[utf8]{inputenc}

\title{Project Proposal Machine Learning}
\author{}
\date{}

\begin{document}
\maketitle

% 
% 
% > 1) Title project 
% > 2) Student member names with student numbers 
% > 3) Application, explain how you get data or simulator, how many examples, how 
% > many inputs, outputs, whether it is an RL, classification, or regression problem. 
% > 4) Methods. Explain the methods that you will compare. Identify the learning 
% > algorithm and the optimization method.

\paragraph{Title project:}
Structured Prediction for Named Entity Recognition

\paragraph{Student member names \& number:}

Joachim Daiber: 2397331\\
Carmen Klaussner: 2401541\\

\paragraph{Application}
We will obtain our training data from CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195}, which contains
data in English and German.Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish. 
Named Entity Recognition is a classification problem in which the goal is to correctly predict the named entity types for
the tokens in a text. Table blala shows the input/output distribution.

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l}
 \textbf{English Data}& Articles & Sentences & Tokens \\ \hline
 Training set &   946       &   14,987        &   203,621     \\
 Development set& 216        &  3,466         &    51,362     \\
 Test set       & 231        &   3,684        &     46,435    \\

 \end{tabular}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l|l}
 \textbf{English Data} & LOC      & MISC     & ORG &PER       \\ \hline
 Training set &   7140       &  3438  & 6321     & 6600    \\
 Development set&  1837       &  922         &  1341    &  1842    \\
 Test set       &  1668       &   702        &   1661    &  1617   \\

 \end{tabular}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l}
 \textbf{German Data} & Articles & Sentences & Tokens \\ \hline
 Training set &     553     &  12,705         &  206,931      \\ 
 Development set&    201     &  3,068         &  51,444       \\
 Test set       &    155     &   3,160        &   51,943      \\

 \end{tabular}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l|l}
 \textbf{German Data} & LOC      & MISC     & ORG &PER       \\        \hline
 Training set &   4363       &   2288        & 2427     & 2773    \\
 Development set&  1181       &   1010        &  1241    &  1401    \\
 Test set       &   1035      &    670       &  773     &  1195   \\

 \end{tabular}
\end{table}

\paragraph{Methods:}
We believe that Structured Prediction provides a flexible and efficient model for Named Entity Recognition. 
The learning algorithm Structured Perceptron with Averaging \cite{Collins:2002:DTM:1118693.1118694}.

% > 6) Chosen programming language 
% > 7) Planning. Describe the different phases and the planning deadlines (until November).
% 
\paragraph{Setup of Experiments:}
We will compare multiple sets of features on the given training and test set, 
while trying to find a good balance between complexity and performance.

\paragraph{Chosen programming language:}
Python with the NumPy package

\paragraph{Planning:}
17-23 Sep.: Data Preparation/Literature Review

24-30 Sep.: Implementation of Learning Algorithm

1-7 Okt.: Implementation of Decoding

8-14 Okt.: Improvement of Features

15-21 Okt.: Evaluation

22-28 Okt.: Paper

29-4 Nov.: Paper

{\small
\bibliographystyle{plain}
\bibliography{ML.bib}
}

\end{document}